# Reproducible Research: Peer Assessment 1
  
## Loading and preprocessing the data
<br>
Data is loaded and preprocessed.
```{r dataclean, echo=TRUE}
activity <- read.csv("./data/activity.csv")
activity$date <- as.Date(activity$date, "%Y-%m-%d")
```
<br>
## What is mean total number of steps taken per day?

The histogram of the total number of steps taken per day.
```{r dailytotal, echo=TRUE}
activity_dailytotal <- aggregate(activity[,c("steps"),drop=FALSE], list(date=activity$date), sum, na.rm=TRUE)
hist(activity_dailytotal$steps, breaks=20, main="", xlab = "Number of steps", ylab = "Frequency")
activity_dailytotalmean <- mean(activity_dailytotal$steps)
activity_dailytotalmedian <- median(activity_dailytotal$steps)
```
The mean and median total number of steps taken per day are `r round(activity_dailytotalmean, digits=2)` and `r activity_dailytotalmedian` respectively.
<br>
## What is the average daily activity pattern?
<br/>
Plot of the average number of steps taken across all days by the 5-minute interval.
```{r intervalmean, echo=TRUE}
activity_intervalmean <- aggregate(activity[,c("steps"),drop=FALSE], list(interval=activity$interval), mean, na.rm=TRUE)
plot(activity_intervalmean$interval, activity_intervalmean$steps, type="l", main = "", xlab = "Interval", ylab = "Number of steps")
activity_maxintervalmean <- activity_intervalmean[activity_intervalmean$steps==max(activity_intervalmean$steps),1]
```
<br>
The maximum number of steps, `r round(max(activity_intervalmean$steps), digits=2)`, occurs in the 5-minute interval that begins at `r sprintf("%02d:%02d", activity_maxintervalmean%/%100, activity_maxintervalmean%%100)`.
## Imputing missing values
<br>
Calculate the number of missing values
```{r missigvalue, echo=TRUE}
activity_countofnasteps <- sum(is.na(activity$steps))
```
There are `r activity_countofnasteps` missing steps values across a number of days/intervals.
<br>
Histogram of the total number of steps taken per day using the **imputed** data.
```{r dailytotal_imputed, echo=TRUE}
activity_imputed <- activity
activity_imputed$steps <- ifelse(is.na(activity_imputed$steps), activity_intervalmean[activity_intervalmean$interval==activity$interval,2], activity$steps)
activity_dailytotal_imputed <- aggregate(activity_imputed[,c("steps"),drop=FALSE], list(date=activity$date), sum, na.rm=TRUE)
hist(activity_dailytotal_imputed$steps, breaks=20, main="", xlab = "Number of steps", ylab = "Frequency")
activity_dailytotalmean_imputed <- mean(activity_dailytotal_imputed$steps)
activity_dailytotalmedian_imputed <- as.integer(median(activity_dailytotal_imputed$steps))
```
The mean and median total number of steps taken per day are `r round(activity_dailytotalmean_imputed, digits=2)` and `r activity_dailytotalmedian_imputed` respectively.
<br>
## Are there differences in activity patterns between weekdays and weekends?
<br>
A series of plot that contains a time series plots of the 5-minute interval and the average number of steps taken across all weekday days or weekend days using the **imputed** data.
```{r intervalmean_imputed, echo=TRUE}
activity_imputed$weekdays <- weekdays(activity_imputed$date)
activity_imputed$weekday <- as.factor(ifelse(activity_imputed$weekdays %in% c("Saturday", "Sunday"), "weekend", "weekday"))
activity_intervalmean_imputed <- aggregate(activity_imputed[,c("steps"),drop=FALSE], list(interval=activity_imputed$interval, weekday=activity_imputed$weekday), mean, na.rm=TRUE)
library(lattice)
p <- xyplot(steps ~ interval | weekday, data = activity_intervalmean_imputed, type = "l", layout = c(1, 2), main = "", xlab = "Interval", ylab = "Number of steps")
print(p)
```